Dynamical Divergence Analysis

### 2017-04-23 analysis-dynamical-MSF-structure.Rmd
bibliografíapaper, varios r0 y dinámica

```{r setup, include = FALSE}
# set chunk options
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r load_packages}
# load packages
library(ggplot2)
library(reshape2)
library(plyr)
library(quantreg)
library(bio3d)

@@

```{r set_my_colors}
# set my colors
myColors = c("blue", "darkgreen","darkgreen", "red", "red", "red") 
names(myColors) <- c("exp", "mut", "no-selection", "medium", "selection", "mut+sel")
```

```{r set_run_variables}
# set significance factors
f95 = 1.96
f99 = 2.6
f999 = 3.3
```

```{r set_file_names}
# set input filenames
## data
reference.fname = file.path(data.dir, paste(family, "_ref.txt", sep = ""))
protein_list.fname = file.path(data.dir, paste(family, "_list.txt", sep = ""))
site.info.fname = file.path(data.dir, paste(p.ref, "_m.da.ca.csv", sep = ""))
min.da.CM.CA.fname = file.path(data.dir, paste(p.ref, "_min.da.CM.ca.csv", sep = ""))

@@

```{r set_reference_protein}
# which is the reference protein?
reference = as.character(read.table(reference.fname, header = F)[1, 1])
if (reference != p.ref) {
  print("Warning: different references")
}
```

We studied the `r family` family. 
We used the ENM `r enm` with R0 = `r R0` to generate mutants.
We used as reference for the ENM the protein with pdb code `r reference`. 
We simulated `r n.mut.p` mutants of this reference for each protein of the family.
The `r family` family is of the type `r type` and consists of the following proteins:

```{r set_experimental_protein_names, comment = ""}
# experimental protein names
protein.exp = read.table(protein_list.fname, header = F, stringsAsFactors = F)[, 1]
protein.exp = protein.exp[protein.exp != reference]
print(protein.exp)
```

@@

# read reference pdb and calculate resno, AA, and CN
reference.pdb.fname = file.path(data.dir,paste(reference, ".pdb", sep = ""))
ref.pdb = read.pdb(reference.pdb.fname)
inds = atom.select(ref.pdb, elety = "CA")
ref.pdb.ca = ref.pdb$atom[inds$atom,]
cmap.ca = cmap(ref.pdb$xyz[inds$xyz], dcut = R0, scut = 0, mask.lower = F)
CN = rowSums(cmap.ca, na.rm = T) - 1 # diag(cmap.ca) returned by cmap is 1!
n.sites = length(CN)
AA = ref.pdb.ca$resid
resno = ref.pdb.ca$resno
pdb.info = data.frame(resno, AA, CN)
names(pdb.info) = c("resno", "AA", paste("CN", as.character(R0), sep = ""))

@@

```{r read_site_data}
# read MSF data
square.dif.MSF.exp = read.csv(square.dif.MSF.exp.fname, header = T)
square.dif.MSF.mut = read.csv(square.dif.MSF.mut.fname, header = T)
square.dif.MSF.medium = read.csv(square.dif.MSF.medium.fname, header = T)

# read data
dri2.exp = read.csv(dri2.exp.fname, header = T)
dri2.mut = read.csv(dri2.mut.fname, header = T)
dri2.medium = read.csv(dri2.medium.fname, header = T)
```

```{r prepare-site-data-MSF}

# prepare the data
## exp
tsquare.dif.MSF.exp = as.data.frame(t(as.matrix(square.dif.MSF.exp)))
tsquare.dif.MSF.exp = cbind(site.info, tsquare.dif.MSF.exp)
siteComparison.exp = melt(tsquare.dif.MSF.exp, id.vars = names(site.info), variable.name = "protein", value.name = "square.dif.MSF")

@@

```{r prepare-site-data-dri2}

# prepare the data
## exp
tdri2.exp = as.data.frame(t(as.matrix(dri2.exp)))
tdri2.exp = cbind(site.info, tdri2.exp)
siteComparison.exp = melt(tdri2.exp, id.vars = names(site.info), variable.name = "protein", value.name = "dri2")

@@

## dat with means and standard errors
d = ddply(dat,c("dataset", "site", "AA", "shell"), function(x) {
                mean.r.SDF = mean(x$r.SDF, na.rm = T)
                ndata = sum(!is.na(x$r.SDF))
                se.r.SDF = sd(x$r.SDF, na.rm = T) / sqrt(ndata)
                data.frame(mean.r.SDF, ndata, se.r.SDF)
})

@@

plot(x = mean.r.square.dif.MSF.exp, 
     y = mean.r.square.dif.MSF.mut, 
     xlab = "exp.mean.r.square.dif.MSF", 
     ylab = "no.selection.mean.r.square.dif.MSF")
abline(0, 1)

plot(x = mean.r.square.dif.MSF.exp, 
     y = mean.r.square.dif.MSF.medium, 
     xlab = "exp.mean.r.square.dif.MSF", 
     ylab = "selection.mean.r.square.dif.MSF")
abline(0, 1)

plot(x = mean.r.square.dif.MSF.mut, 
     y = mean.r.square.dif.MSF.medium, 
     xlab = "no.selection.mean.r.square.dif.MSF", 
     ylab = "selection.mean.r.square.dif.MSF")
abline(0, 1)
```

```{r mean-z.RSD-profiles}

@@

# dat with means and standard errors
d = ddply(dat,c("dataset", "site", "AA", "shell"), function(x) {
                mean.z.RSD = mean(x$z.RSD, na.rm = T)
                ndata = sum(!is.na(x$z.RSD))
                se.z.RSD = sd(x$z.RSD, na.rm = T) / sqrt(ndata)
                data.frame(mean.z.RSD, ndata, se.z.RSD)
})

@@

plot(x = mean.r.square.dif.MSF.exp, 
     y = mean.r.dri2.exp, 
     xlab = "exp.mean.r.square.dif.MSF", 
     ylab = "exp.mean.r.dri2")
abline(0, 1)

plot(x = mean.r.square.dif.MSF.medium, 
     y = mean.r.dri2.medium, 
     xlab = "selection.mean.r.square.dif.MSF", 
     ylab = "selection.mean.r.dri2")
abline(0, 1)

plot(x = mean.r.square.dif.MSF.mut, 
     y = mean.r.dri2.mut, 
     xlab = "no.selection.mean.r.square.dif.MSF", 
     ylab = "no.selection.mean.r.dri2")
abline(0, 1)

>>>>

```{r setup, include = FALSE}
# set chunk options
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```
```{r load-packages}
# load packages
library(ggplot2)
library(reshape2)
library(plyr)
library(quantreg)
library(bio3d)

@@

```{r set-my-colors}
# set my colors
myColors = c("blue", "darkgreen","darkgreen", "red", "red", "red") 
names(myColors) <- c("exp", "mut", "no-selection", "medium", "selection", "mut+sel")
```

```{r set-run-variables}
# general parameters
tolerance = 1e-10

# set significance factors
f95 = 1.96
f99 = 2.6
f999 = 3.3
```

```{r set-file-names}
# set input filenames
## data
reference.fname = file.path(data.dir, paste(family, "_ref.txt", sep = ""))
protein_list.fname = file.path(data.dir, paste(family, "_list.txt", sep = ""))
site.info.fname = file.path(data.dir, paste(p.ref, "_m.da.ca.csv", sep = ""))
min.da.CM.CA.fname = file.path(data.dir, paste(p.ref, "_min.da.CM.ca.csv", sep = ""))

@@

```
```{r set-reference-protein}
# which is the reference protein?
reference = as.character(read.table(reference.fname, header = F)[1, 1])
if (reference != p.ref) {
  print("Warning: different references")
}
```
We studied the `r family` family. 
We used the ENM `r enm` with R0 = `r R0` to generate mutants.
We used as reference for the ENM the protein with pdb code `r reference`. 
We simulated `r n.mut.p` mutants of this reference for each protein of the family.
The `r family` family is of the type `r type` and consists of the following proteins:

```{r set-experimental-protein-names, comment = ""}
# experimental protein names
protein.exp = read.table(protein_list.fname, header = F, stringsAsFactors = F)[, 1]
protein.exp = protein.exp[protein.exp != reference]
print(protein.exp)
```

@@

# read reference pdb and calculate resno, AA, and CN
reference.pdb.fname = file.path(data.dir,paste(reference, ".pdb", sep = ""))
ref.pdb = read.pdb(reference.pdb.fname)
inds = atom.select(ref.pdb, elety = "CA")
ref.pdb.ca = ref.pdb$atom[inds$atom,]
cmap.ca = cmap(ref.pdb$xyz[inds$xyz], dcut = R0, scut = 0, mask.lower = F)
r.p.ref = ref.pdb$xyz[inds$xyz]
CN = rowSums(cmap.ca, na.rm = T) - 1 # diag(cmap.ca) returned by cmap is 1!
n.sites = length(CN)
AA = ref.pdb.ca$resid
resno = ref.pdb.ca$resno
pdb.info = data.frame(resno, AA, CN)
names(pdb.info) = c("resno", "AA", paste("CN", as.character(R0), sep = ""))

@@

```{r read-site-data}
# read MSF data
square.dif.MSF.exp = read.csv(square.dif.MSF.exp.fname, header = T)
square.dif.MSF.mut = read.csv(square.dif.MSF.mut.fname, header = T)
square.dif.MSF.medium = read.csv(square.dif.MSF.medium.fname, header = T)

# read dri2 data
dri2.exp = read.csv(dri2.exp.fname, header = T)
dri2.mut = read.csv(dri2.mut.fname, header = T)
dri2.medium = read.csv(dri2.medium.fname, header = T)
```

```{r MSF-p.ref}
# calculate MSF (Mean Square Fluctuation)

## calculate ENM
ENMK.p.ref = CalculateENMKeff(matrix(r.p.ref, nrow = 3), 
                              seq(1:n.sites), 
                              c(), 
                              R0, 
                              tolerance, 
                              K.analysis)
## get cov matrix
cov.p.ref = ENMK.p.ref$cov

## get the diagonal of the cov matrix
diag.p.ref = diag(cov.p.ref)

## calculate the factor to split the diagonal
factor = sort(rep(seq(1:n.sites), 3))

## split the diagonal
s.diag.p.ref = split(diag.p.ref, factor)

## create a matrix to save the data
MSF.p.ref = matrix(nrow = 1, ncol = n.sites)

## start a loop to calculate sums for each site
for (i in (1:n.sites)) {
  MSF.p.ref[, i] = sum(unlist(s.diag.p.ref[i]), use.names = F)
}

```{r prepare-site-data-MSF}
# prepare the data
## exp
tsquare.dif.MSF.exp = as.data.frame(t(as.matrix(square.dif.MSF.exp)))
tsquare.dif.MSF.exp = cbind(site.info, tsquare.dif.MSF.exp)
siteComparison.exp = melt(tsquare.dif.MSF.exp, id.vars = names(site.info), variable.name = "protein", value.name = "square.dif.MSF")

@@

```{r prepare-site-data-dri2}
# prepare the data
## exp
tdri2.exp = as.data.frame(t(as.matrix(dri2.exp)))
tdri2.exp = cbind(site.info, tdri2.exp)
siteComparison.exp = melt(tdri2.exp, id.vars = names(site.info), variable.name = "protein", value.name = "dri2")

@@

## dat with means and standard errors
d = ddply(dat,c("dataset", "site", "AA", "shell"), function(x) {
                mean.r.SDF = mean(x$r.SDF, na.rm = T)
                n.data = sum(!is.na(x$r.SDF))
                se.r.SDF = sd(x$r.SDF, na.rm = T) / sqrt(n.data)
                data.frame(mean.r.SDF, n.data, se.r.SDF)
})

@@

plot(x = mean.r.square.dif.MSF.exp, 
     y = mean.r.square.dif.MSF.mut, 
     xlab = "<r.square.dif.MSF> experimental", 
     ylab = "<r.square.dif.MSF> no-selection")
abline(0, 1)

plot(x = mean.r.square.dif.MSF.exp, 
     y = mean.r.square.dif.MSF.medium, 
     xlab = "<r.square.dif.MSF> experimental", 
     ylab = "<r.square.dif.MSF> selection")
abline(0, 1)

plot(x = mean.r.square.dif.MSF.mut, 
     y = mean.r.square.dif.MSF.medium, 
     xlab = "<r.square.dif.MSF> no-selection", 
     ylab = "<r.square.dif.MSF> selection")
abline(0, 1)
```

```{r comparison-with-MSF.p.ref}

# rank MSF.p.ref
r.MSF.p.ref = rank(MSF.p.ref)

# calculate correlations
r.exp = cor(mean.r.square.dif.MSF.exp, r.MSF.p.ref)
r.mut = cor(mean.r.square.dif.MSF.mut, r.MSF.p.ref)
r.medium = cor(mean.r.square.dif.MSF.medium, r.MSF.p.ref)

cor.df = data.frame("comparison" = c("<r.square.dif.MSF> experimental vs r.MSF.p.ref", "<r.square.dif.MSF> no-selection vs r.MSF.p.ref", "<r.square.dif.MSF> selection vs r.MSF.p.ref"),
                    "R" = c(r.exp, r.mut, r.medium))

plot(x = mean.r.square.dif.MSF.exp, 
     y = r.MSF.p.ref, 
     xlab = "<r.square.dif.MSF> experimental", 
     ylab = "r.MSF.p.ref")
abline(0, 1)

plot(x = mean.r.square.dif.MSF.mut, 
     y = r.MSF.p.ref, 
     xlab = "<r.square.dif.MSF> no-selection", 
     ylab = "r.MSF.p.ref")
abline(0, 1)

plot(x = mean.r.square.dif.MSF.medium, 
     y = r.MSF.p.ref, 
     xlab = "<r.square.dif.MSF> selection", 
     ylab = "r.MSF.p.ref")
abline(0, 1)
```
```{r mean-z.RSD-profiles}

@@

# dat with means and standard errors
d = ddply(dat,c("dataset", "site", "AA", "shell"), function(x) {
                mean.z.RSD = mean(x$z.RSD, na.rm = T)
                n.data = sum(!is.na(x$z.RSD))
                se.z.RSD = sd(x$z.RSD, na.rm = T) / sqrt(n.data)
                data.frame(mean.z.RSD, n.data, se.z.RSD)
})

@@

plot(x = mean.r.square.dif.MSF.exp, 
     y = mean.r.dri2.exp, 
     xlab = "r<r.square.dif.MSF> experimental", 
     ylab = "r<dri2> experimental")
abline(0, 1)

plot(x = mean.r.square.dif.MSF.mut, 
     y = mean.r.dri2.mut, 
     xlab = "r<r.square.dif.MSF> no-selection", 
     ylab = "r<dri2> no-selection")
abline(0, 1)

plot(x = mean.r.square.dif.MSF.medium, 
     y = mean.r.dri2.medium, 
     xlab = "r<r.square.dif.MSF> selection", 
     ylab = "r<dri2> selection")
abline(0, 1)

### 2017-04-23 analysis-dynamical-nH.Rmd
bibliografíapaper, varios r0 y dinámica

---
title: "Structural Divergence Analysis"
author: "Maria Laura Marcos, Julian Echave"
output: html_document
---

@@

```{r set_my_colors}
# set my colors
myColors = c("blue", "darkgreen", "red", "red") 
names(myColors) <- c("exp", "mut", "medium", "mut+sel")
```

```{r set_run_variables}
# set significance factors
f95 = 1.96
f99 = 2.6
f999 = 3.3
```

```{r set_file_names}
# set input filenames
## data
reference.fname = file.path(data.dir, paste(family, "_ref.txt", sep = ""))
protein_list.fname = file.path(data.dir, paste(family, "_list.txt", sep = ""))

## profiles to compare
### nH
nH.exp.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.exp.nH.csv", sep = ""))
nH.medium.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_medium.sel_K.analysis_Keff_out_m.theo.nH.csv", sep = ""))
nH.mut.fname = file.path(data.dir,paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.theo.nH.csv", sep = ""))

```

```{r set_reference_protein}
# which is the reference protein?
reference = as.character(read.table(reference.fname, header = F)[1, 1])
if (reference != p.ref) {
  print("Warning: different references")
}
```

We studied the `r family` family. 
We used the ENM `r enm` with R0 = `r R0` to generate mutants.
We used as reference for the ENM the protein with pdb code `r reference`. 
We simulated `r n.mut.p` mutants of this reference for each protein of the family.
The `r family` family is of the type `r type` and consists of the following proteins:

```{r set_experimental_protein_names, comment = ""}
# experimental protein names
protein.exp = read.table(protein_list.fname, header = F, stringsAsFactors = F)[, 1]
protein.exp = protein.exp[protein.exp != reference]
print(protein.exp)
```

```{r prepare_mode_data}
# prepare the data
nH.exp = read.csv(nH.exp.fname, header = T)
nH.mut = read.csv(nH.mut.fname, header = T)
nH.medium = read.csv(nH.medium.fname, header = T)
```

### Dynamical divergence along normal modes

In this section we consider dynamical divergence using normal modes coordinates. We calculated mean nH and we z-normalized profiles.

@@

# scale profiles
mean.nH.exp = scale(mean.nH.exp[which(!is.na(mean.nH.exp))])
mean.nH.mut = scale(mean.nH.mut[which(!is.na(mean.nH.mut))])
mean.nH.medium = scale(mean.nH.medium[which(!is.na(mean.nH.medium))])

plot(mean.nH.exp, ylab = "<nH>", xlab = "mode", col = "blue")
points(mean.nH.mut, col = "green")
points(mean.nH.medium, col = "red")

@@

```{r vs-mean-profiles}

# plot exp vs simulated profiles 
par(pty = "s")

plot(x = mean.nH.exp, 
     y = mean.nH.mut, 
     xlab = "exp.mean.nH", 
     ylab = "no.selection.mean.nH")
abline(0, 1)

plot(x = mean.nH.exp, 
     y = mean.nH.medium, 
     xlab = "exp.mean.nH", 
     ylab = "selection.mean.nH")
abline(0, 1)

plot(x = mean.nH.mut, 
     y = mean.nH.medium, 
     xlab = "no.selection.mean.nH", 
     ylab = "selection.mean.nH")
abline(0, 1)
```

@@

# get first 100 modes and re-scale
mean.nH.exp = scale(mean.nH.exp[1:100])
mean.nH.mut = scale(mean.nH.mut[1:100])
mean.nH.medium = scale(mean.nH.medium[1:100])

plot(mean.nH.exp, ylab = "<nH>", xlab = "mode", col = "blue")
points(mean.nH.mut, col = "green")
points(mean.nH.medium, col = "red")

@@

cor.df = data.frame("comparison" = c("exp vs no-selection", "exp vs selection", "no-selection vs selection"),
                    "R" = c(r.exp.mut, r.exp.medium, r.mut.medium))
knitr::kable(cor.df, digits = 2)

# plot exp vs simulated profiles 
par(pty = "s")

plot(x = mean.nH.exp, 
     y = mean.nH.mut, 
     xlab = "exp.mean.nH", 
     ylab = "no.selection.mean.nH")
abline(0, 1)

plot(x = mean.nH.exp, 
     y = mean.nH.medium, 
     xlab = "exp.mean.nH", 
     ylab = "selection.mean.nH")
abline(0, 1)

plot(x = mean.nH.mut, 
     y = mean.nH.medium, 
     xlab = "no.selection.mean.nH", 
     ylab = "selection.mean.nH")
abline(0, 1)
```

>>>>

---
title: "Dynamical Divergence Analysis"
author: "Maria Laura Marcos, Julian Echave"
output: html_document
---

@@

```{r set-filenames}
# set input filenames
## data
reference.fname = file.path(data.dir, paste(family, "_ref.txt", sep = ""))
protein_list.fname = file.path(data.dir, paste(family, "_list.txt", sep = ""))

## profiles to compare
### nH
nH.exp.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.exp.nH.csv", sep = ""))
nH.medium.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_medium.sel_K.analysis_Keff_out_m.theo.nH.csv", sep = ""))
nH.mut.fname = file.path(data.dir,paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.theo.nH.csv", sep = ""))
```

```{r set_reference_protein}
# which is the reference protein?
reference = as.character(read.table(reference.fname, header = F)[1, 1])
if (reference != p.ref) {
  print("Warning: different references")
}
```
We studied the `r family` family. 
We used the ENM `r enm` with R0 = `r R0` to generate mutants.
We used as reference for the ENM the protein with pdb code `r reference`. 
We simulated `r n.mut.p` mutants of this reference for each protein of the family.
The `r family` family is of the type `r type` and consists of the following proteins:

```{r set-experimental-protein-names, comment = ""}
# experimental protein names
protein.exp = read.table(protein_list.fname, header = F, stringsAsFactors = F)[, 1]
protein.exp = protein.exp[protein.exp != reference]
print(protein.exp)
```
```{r prepare-mode-data}
# prepare the data
nH.exp = read.csv(nH.exp.fname, header = T)
nH.mut = read.csv(nH.mut.fname, header = T)
nH.medium = read.csv(nH.medium.fname, header = T)
```
### Dynamical divergence along normal modes

In this section we consider dynamical divergence using normal modes coordinates. We calculated <nH> and we z-normalized profiles.

@@

# scale profiles
mean.nH.exp = vscale(mean.nH.exp[which(!is.na(mean.nH.exp))])
mean.nH.mut = vscale(mean.nH.mut[which(!is.na(mean.nH.mut))])
mean.nH.medium = vscale(mean.nH.medium[which(!is.na(mean.nH.medium))])

plot(mean.nH.exp, ylab = "<z.nH>", xlab = "mode", col = "blue")
points(mean.nH.mut, col = "green")
points(mean.nH.medium, col = "red")

@@

```{r vs-mean-plots}

# plot comparative profiles 
par(pty = "s")

plot(x = mean.nH.exp, 
     y = mean.nH.mut, 
     xlab = "<z.nH> experimental", 
     ylab = "<z.nH> no-selection")
abline(0, 1)

plot(x = mean.nH.exp, 
     y = mean.nH.medium, 
     xlab = "<z.nH> experimental", 
     ylab = "<z.nH> selection")
abline(0, 1)

plot(x = mean.nH.mut, 
     y = mean.nH.medium, 
     xlab = "<z.nH> no-selection", 
     ylab = "<z.nH> selection")
abline(0, 1)
```

@@

# get first 100 modes and re-scale
mean.nH.exp = vscale(mean.nH.exp[1:100])
mean.nH.mut = vscale(mean.nH.mut[1:100])
mean.nH.medium = vscale(mean.nH.medium[1:100])

plot(mean.nH.exp, ylab = "<z.nH>", xlab = "mode", col = "blue")
points(mean.nH.mut, col = "green")
points(mean.nH.medium, col = "red")

@@

```{r vs-mean-plots-100-modes}

# plot comparative profiles 
plot(x = mean.nH.exp, 
     y = mean.nH.mut, 
     xlab = "<z.nH> experimental", 
     ylab = "<z.nH> no-selection")
abline(0, 1)

plot(x = mean.nH.exp, 
     y = mean.nH.medium, 
     xlab = "<z.nH> experimental", 
     ylab = "<z.nH> selection")
abline(0, 1)

plot(x = mean.nH.mut, 
     y = mean.nH.medium, 
     xlab = "<z.nH> no-selection", 
     ylab = "<z.nH> selection")
abline(0, 1)
```
```{r comparison-of-means-20-modes}

# get first 100 modes and re-scale
mean.nH.exp = vscale(mean.nH.exp[1:20])
mean.nH.mut = vscale(mean.nH.mut[1:20])
mean.nH.medium = vscale(mean.nH.medium[1:20])

plot(mean.nH.exp, ylab = "<z.nH>", xlab = "mode", col = "blue")
points(mean.nH.mut, col = "green")
points(mean.nH.medium, col = "red")

```{r vs-mean-plots-20-modes}

# plot comparative profiles 
plot(x = mean.nH.exp, 
     y = mean.nH.mut, 
     xlab = "<z.nH> experimental", 
     ylab = "<z.nH> no-selection")
abline(0, 1)

plot(x = mean.nH.exp, 
     y = mean.nH.medium, 
     xlab = "<z.nH> experimental", 
     ylab = "<z.nH> selection")
abline(0, 1)

plot(x = mean.nH.mut, 
     y = mean.nH.medium, 
     xlab = "<z.nH> no-selection", 
     ylab = "<z.nH> selection")
abline(0, 1)
```

### 2017-04-23 analysis-dynamical-nR.Rmd
bibliografíapaper, varios r0 y dinámica

```{r set_my_colors}
# set my colors
myColors = c("blue", "darkgreen", "red", "red") 
names(myColors) <- c("exp", "mut", "medium", "mut+sel")
```

```{r set_run_variables}
# set significance factors
f95 = 1.96
f99 = 2.6
f999 = 3.3
```

```{r set_file_names}
# set input filenames
## data
reference.fname = file.path(data.dir, paste(family, "_ref.txt", sep = ""))
protein_list.fname = file.path(data.dir, paste(family, "_list.txt", sep = ""))

## profiles to compare
### nR
nR.exp.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.exp.nR.csv", sep = ""))
nR.medium.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_medium.sel_K.analysis_Keff_out_m.theo.nR.csv", sep = ""))
nR.mut.fname = file.path(data.dir,paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.theo.nR.csv", sep = ""))

```

```{r set_reference_protein}
# which is the reference protein?
reference = as.character(read.table(reference.fname, header = F)[1, 1])
if (reference != p.ref) {
  print("Warning: different references")
}
```

We studied the `r family` family. 
We used the ENM `r enm` with R0 = `r R0` to generate mutants.
We used as reference for the ENM the protein with pdb code `r reference`. 
We simulated `r n.mut.p` mutants of this reference for each protein of the family.
The `r family` family is of the type `r type` and consists of the following proteins:

```{r set_experimental_protein_names, comment = ""}
# experimental protein names
protein.exp = read.table(protein_list.fname, header = F, stringsAsFactors = F)[, 1]
protein.exp = protein.exp[protein.exp != reference]
print(protein.exp)
```

```{r prepare_mode_data}
# prepare the data
nR.exp = read.csv(nR.exp.fname, header = T)
nR.mut = read.csv(nR.mut.fname, header = T)
nR.medium = read.csv(nR.medium.fname, header = T)
```

### Dynamical divergence along normal modes

In this section we consider dynamical divergence using normal modes coordinates. We calculated mean nR and we z-normalized profiles.

@@

# scale profiles
mean.nR.exp = scale(mean.nR.exp[which(!is.na(mean.nR.exp))])
mean.nR.mut = scale(mean.nR.mut[which(!is.na(mean.nR.mut))])
mean.nR.medium = scale(mean.nR.medium[which(!is.na(mean.nR.medium))])

plot(mean.nR.exp, ylab = "<nR>", xlab = "mode", col = "blue")
points(mean.nR.mut, col = "green")
points(mean.nR.medium, col = "red")

@@

```{r vs-mean-profiles}

# plot exp vs simulated profiles 
par(pty = "s")

plot(x = mean.nR.exp, 
     y = mean.nR.mut, 
     xlab = "exp.mean.nR", 
     ylab = "no.selection.mean.nR")
abline(0, 1)

plot(x = mean.nR.exp, 
     y = mean.nR.medium, 
     xlab = "exp.mean.nR", 
     ylab = "selection.mean.nR")
abline(0, 1)

plot(x = mean.nR.mut, 
     y = mean.nR.medium, 
     xlab = "no.selection.mean.nR", 
     ylab = "selection.mean.nR")
abline(0, 1)
```

@@

# get first 100 modes and re-scale
mean.nR.exp = scale(mean.nR.exp[1:100])
mean.nR.mut = scale(mean.nR.mut[1:100])
mean.nR.medium = scale(mean.nR.medium[1:100])

plot(mean.nR.exp, ylab = "<nR>", xlab = "mode", col = "blue")
points(mean.nR.mut, col = "green")
points(mean.nR.medium, col = "red")

@@

cor.df = data.frame("comparison" = c("exp vs no-selection", "exp vs selection", "no-selection vs selection"),
                    "R" = c(r.exp.mut, r.exp.medium, r.mut.medium))
knitr::kable(cor.df, digits = 2)

# plot exp vs simulated profiles 
par(pty = "s")

plot(x = mean.nR.exp, 
     y = mean.nR.mut, 
     xlab = "exp.mean.nR", 
     ylab = "no.selection.mean.nR")
abline(0, 1)

plot(x = mean.nR.exp, 
     y = mean.nR.medium, 
     xlab = "exp.mean.nR", 
     ylab = "selection.mean.nR")
abline(0, 1)

plot(x = mean.nR.mut, 
     y = mean.nR.medium, 
     xlab = "no.selection.mean.nR", 
     ylab = "selection.mean.nR")
abline(0, 1)
```

>>>>

```{r set-filenames}
# set input filenames
## data
reference.fname = file.path(data.dir, paste(family, "_ref.txt", sep = ""))
protein_list.fname = file.path(data.dir, paste(family, "_list.txt", sep = ""))

## profiles to compare
### nR
nR.exp.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.exp.nR.csv", sep = ""))
nR.medium.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_medium.sel_K.analysis_Keff_out_m.theo.nR.csv", sep = ""))
nR.mut.fname = file.path(data.dir,paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.theo.nR.csv", sep = ""))
```

```{r set_reference_protein}
# which is the reference protein?
reference = as.character(read.table(reference.fname, header = F)[1, 1])
if (reference != p.ref) {
  print("Warning: different references")
}
```
We studied the `r family` family. 
We used the ENM `r enm` with R0 = `r R0` to generate mutants.
We used as reference for the ENM the protein with pdb code `r reference`. 
We simulated `r n.mut.p` mutants of this reference for each protein of the family.
The `r family` family is of the type `r type` and consists of the following proteins:

```{r set-experimental-protein-names, comment = ""}
# experimental protein names
protein.exp = read.table(protein_list.fname, header = F, stringsAsFactors = F)[, 1]
protein.exp = protein.exp[protein.exp != reference]
print(protein.exp)
```
```{r prepare-mode-data}
# prepare the data
nR.exp = read.csv(nR.exp.fname, header = T)
nR.mut = read.csv(nR.mut.fname, header = T)
nR.medium = read.csv(nR.medium.fname, header = T)
```
### Dynamical divergence along normal modes

In this section we consider dynamical divergence using normal modes coordinates. We calculated <nR> and we z-normalized profiles.

@@

# scale profiles
mean.nR.exp = vscale(mean.nR.exp[which(!is.na(mean.nR.exp))])
mean.nR.mut = vscale(mean.nR.mut[which(!is.na(mean.nR.mut))])
mean.nR.medium = vscale(mean.nR.medium[which(!is.na(mean.nR.medium))])

plot(mean.nR.exp, ylab = "<z.nR>", xlab = "mode", col = "blue")
points(mean.nR.mut, col = "green")
points(mean.nR.medium, col = "red")

@@

```{r vs-mean-plots}

# plot comparative profiles 
par(pty = "s")

plot(x = mean.nR.exp, 
     y = mean.nR.mut, 
     xlab = "<z.nR> experimental", 
     ylab = "<z.nR> no-selection")
abline(0, 1)

plot(x = mean.nR.exp, 
     y = mean.nR.medium, 
     xlab = "<z.nR> experimental", 
     ylab = "<z.nR> selection")
abline(0, 1)

plot(x = mean.nR.mut, 
     y = mean.nR.medium, 
     xlab = "<z.nR> no-selection", 
     ylab = "<z.nR> selection")
abline(0, 1)
```

@@

# get first 100 modes and re-scale
mean.nR.exp = vscale(mean.nR.exp[1:100])
mean.nR.mut = vscale(mean.nR.mut[1:100])
mean.nR.medium = vscale(mean.nR.medium[1:100])

plot(mean.nR.exp, ylab = "<z.nR>", xlab = "mode", col = "blue")
points(mean.nR.mut, col = "green")
points(mean.nR.medium, col = "red")

@@

```{r vs-mean-plots-100-modes}

# plot comparative profiles 
plot(x = mean.nR.exp, 
     y = mean.nR.mut, 
     xlab = "<z.nR> experimental", 
     ylab = "<z.nR> no-selection")
abline(0, 1)

plot(x = mean.nR.exp, 
     y = mean.nR.medium, 
     xlab = "<z.nR> experimental", 
     ylab = "<z.nR> selection")
abline(0, 1)

plot(x = mean.nR.mut, 
     y = mean.nR.medium, 
     xlab = "<z.nR> no-selection", 
     ylab = "<z.nR> selection")
abline(0, 1)
```
```{r comparison-of-means-20-modes}

# get first 100 modes and re-scale
mean.nR.exp = vscale(mean.nR.exp[1:20])
mean.nR.mut = vscale(mean.nR.mut[1:20])
mean.nR.medium = vscale(mean.nR.medium[1:20])

plot(mean.nR.exp, ylab = "<z.nR>", xlab = "mode", col = "blue")
points(mean.nR.mut, col = "green")
points(mean.nR.medium, col = "red")

```{r vs-mean-plots-20-modes}

# plot comparative profiles 
plot(x = mean.nR.exp, 
     y = mean.nR.mut, 
     xlab = "<z.nR> experimental", 
     ylab = "<z.nR> no-selection")
abline(0, 1)

plot(x = mean.nR.exp, 
     y = mean.nR.medium, 
     xlab = "<z.nR> experimental", 
     ylab = "<z.nR> selection")
abline(0, 1)

plot(x = mean.nR.mut, 
     y = mean.nR.medium, 
     xlab = "<z.nR> no-selection", 
     ylab = "<z.nR> selection")
abline(0, 1)
```
